Stack Safety as a Security Property

### 2021-03-18 CheriApproaches.tex
Expanded Cheri writeup

This document deals with questions of the division between machine state and policy state.
In existing models of our security properties, the state is divided into a machine state,
which is a mapping from addresses and registers to data, and a policy state, which may be
any arbitrary data. A machine state can step on its own, or in conjunction with a policy
state; when states step together the machine state component always steps as it would on
its own. A policy state may only cause the step to fail outright, never change the result.

>>>>

\section{A Stack-safe Calling Convention for Capabilities}

This is Skorstengaard's calling convention from the worse of their papers, with a
difference in that we use a single piece of code to handle the ``landing pad'' of a return.
We must combine this with descriptions of how to adjust out model to support them. At the
high level, we need to accept that their threat model is slightly incompatible with ours:
adversarial code, defined as code whose dynamic behavior does not obey simple reasonableness
properties, is exempt from the LSE and WBCF protections in its own calls. For instance, such
a function might make a call in such a way that the return skips over it and returns straight
to its caller, violating WBCF. Or it might create capabilities on its own data and pass them
along, violating LSE. This latter piece of unreasonableness is easy to handle: we add a clause
to integrity to assume that the caller is reasonable. Equivalent adjustments to confidentiality
might be trickier, in that we should acknowledge the possibility of unreasonableness in either
the callee or the caller. But the violations of WBCF are much more problematic, because our
model of counting depths assumes that callers are never ``skipped'' by a multi-level return.

\subsection{Updating Stack Safety to Support Unreasonable Functions}

The solution given in StkTokens is, essentially, that unreasonable functions don't get to
make ``real'' (protected) calls. Rather, they perform all the operations involved in a call,
then jump into another function at its entry point. That function may then jump back -- which
is not a ``real'' return -- or do an actual return to the original caller. The equivalent to
this is in our model is for us to, on a call-annotated instruction, check that the current
function is reasonable. If it is not, we do not update the depth, and given the callee an
unsealed piece of stack to work with. Similarly, on return, we only update the context
if we are returning to a reasonable function. That requires examining the program counter
that we're returning to. The problem with this whole approach is that it doesn't play nice
with the initialization side of confidentiality. The adversary gets to initialize its
callee's frame before fake-calling it. A reasonable callee will not read those initialized
variables, anyway, but the property will not capture that the protection is needed. We
could potentially treat this as a special kind of call, obeying confidentiality but not
integrity or WBCF, which gets us closer to the next approach.

The more complex approach is to accept that unreasonable callers might get skipped or
accessed, but otherwise model their calls and returns normally. That means extending
our model with support for dynamic returns, which are then forbidden by WBCF in the case
of reasonable callers. Specifically, the {\sc Ret} annotation is now parameterized by
a function that takes a machine state and a context and returns an integer depth: the
return target. Instead of only popping the top depth, we pop the depth
we're returning to and every one above it, by the same mechanisms as in a single return.
In the context of PIPE, where we always guarantee single returns, the function can simply
subtract 1 from the current depth, as before. We will discuss how to implement this
function for Cheri once we cover the calling convention.

Now most of our properties are unaffected by this, even Return Integrity, which after
all talks about what happens {\it when} we return to a caller. Then we simply need a second
property, which I propose to call Return Guarantee, that if a caller is reasonable it
will see a return {\it before} any caller with a lower depth does.

\subsection{Calling Convention Details}

First let's discuss our assumptions about the capability system, starting with basic
capabilities.

A {\em capability} is composed of a {\em base}, {\em offset}, and {\em bound}
Each word of memory carries a 2-bit tag, the first bit indicated whether it is a
{\em valid} capability and the second whether that capability is {\em local}
A local capability can never be stored in memory outside of the stack. A load or store
failstops if the capability used as its pointer does not have its offset within its
bounds. A load or store using a non-capability pointer uses the bounds of one of the
following special capabilities.

There are three special capability registers. The {\em default data capability} ({\tt \$ddc}),
is applied to reads and writes through non-capability pointers. The
{\em program counter capability} ({\tt \$pcc}) is applied to instruction fetches.
The {\em stack capability} ({\tt spc}) takes the place of the stack pointer and is used by
compiler generated code. Basic capabilities are supported by the following instructions:

\begin{tabular}{l l l}
  {\tt st}      & Data Store       & Stores data, clearing any capability bits \\
  {\tt ld}      & Data Load        & Loads data, clearing any capability bits \\
  {\tt cst}     & Capability Store & Stores a capability, but failstops rather than store
                                     a local one outside the stack \\
  {\tt cld}     & Capability Load  & Loads a capability from memory \\
  {\tt incbase} & Increment Base   & Increases the base of a capability by one and
                                     decrements the offset \\
  {\tt decbnd}  & Decrement Bound  & Decreases the bound of a capability by one \\
\end{tabular}

We will need to use {\em object capabilities}. Object capabilities come in twos:
a code pointer and a memory capability. They are sealed with a matching identifier and
can only be unsealed by invoking them together, simultaneously jumping to the code pointer
and unsealing the memory capability. This way we ensure that the memory capability is
only used by trusted code. These are supported by the following instructions:

\begin{tabular}{l l l}
  {\tt so} & Seal Object   & Takes a pair of capabilities and
                                  seals them with matching identifiers \\
  {\tt io} & Invoke Object & Takes matching sealed capabilities \(c\) and
                                  \(p\), jumps to \(p\), and unseals \(c\) \\
\end{tabular}

Finally, we will use monotone capabilities.

\subsection{The calling convention}

From here on out, assume all capabilities are local. They can live in registers and stack
memory only. In the interest of corresponding to out model, the {\tt \$ddc} is permissive
and allows access to everything that isn't stack or code, and the {\tt \$pcc} allows access
only to our code.

{\bf Initially:} The {\tt \$stc} encompasses the full stack in its bounds. There is a
special bit of code that has no function identifier that serves as a landing pad for
functions, that takes a capability on a two-word block, loads the first word to {\tt \$stc}
and jumps to the second. The jump is annotated as a return by the compiler.

{\bf Caller:} To make a call, the caller allocates two extra words on the stack. It copies
the current {\tt \$stc} and the address of the instruction after the call into these. Then
it creates a capability restricted to this object and links it as an object capability with
the address of the landing pad. The object capability is stored in a register. Then the
caller increments the base of the {\tt \$stc} until its offset is zero, which will prevent
access to its frame. Finally, the caller jumps to the call target -- this instruction is
annotated as a call by the compiler.

After a return, the caller immediately clears all non-return-value registers and the stack
above the {\tt \$stc}'s offset. This prevents a copy of the object capability from being
stashed somewhere for a multiple return attack.

{\bf Callee:} On entry, the callee allocates memory by increasing the {\tt \$stc} and stores
the object capability it was passed. To return, it invokes that capability, executing the
landing pad and unsealing the data portion of the object capability. At the end of the
landing pad sequence, there is a jump instruction, which is annotated as a return. As
discussed above, to dynamically determine the target of the return, we must look at the
machine state and contet: the data portion of the return capability points to the activation
record of the caller, which is labeled {\it Passed(d)} for some depth \(d\) in the context.
(That is, it lies with the frame of the activation at depth \(d\).) The return target is
therefore \(d\). An intuitive way of describing this is that on return, we unseal layers of
calls until the activation record we're using to return has been unsealed.

\subsection{Analysis}

This calling convention prevents the callee from accessing the caller's state. The caller
guarantees that there is no additional object capability floating around somewhere that
the callee could access, and the {\tt \$stc} has its bounds shrunk to exclude the caller's
frame. It can only be restored with the object invocation.

What it doesn't necessarily do is prevent the callee from reading its frame. An adversarial
caller could write above its stack pointer before the call and there would be no protection.
So we have ``sealed confidentiality'' instead of normal confidentiality that includes all
uninitialized memory.

Luckily, if we extend our model with the ``uninitialized capabilities'' of Georges et. al.
we can once again enforce the confidentiality of uninitializes memory, as well. Then
we have full local state encapsulation.

The generic return code being shared between all functions is a little unfortunate, as it
puts a hole in control separation. We'll need to describe the correct behavior carefully
and adjust the property.

We may have a problem with returns, in that a callee could potentially pass along its
return object to a further callee, which could then bypass the first and return straight
to the original caller. This appears to be a hazard of our attacker model.

The caller would be happy enough, as its state would have been protected,
but it messes up our whole model of depths and counting
of returns. And the callee might rely on getting returned to.

\section{Well-formedness, Reasonableness, and PIPE}

As discussed above, Skortengaard et. al. impose static well-formedness conditions on
all of their code, as well as certain

\section{The Question of Policy States}

This half of the document deals with questions of the division between machine state and
policy state. In existing models of our security properties, the state is divided into a
machine state, which is a mapping from addresses and registers to data, and a policy state,
which may be any arbitrary data. A machine state can step on its own, or in conjunction
with a policy state; when states step together the machine state component always steps
as it would on its own. A policy state may only cause the step to fail outright, never
change the result.
